---
title: "Workout 1"
author: "Ethan M."
date: "October 2, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) File Structure

Following the commands outlined in the assignment I created the necessary folders: (I know the document says not to use `eval = FALSE` but in order to display the commands I used in terminal I have to)
```{r, eval = FALSE}
cd stat133-hws-fall18
mkdir workout1
cd workout1
touch README.md
mkdir data
mkdir code
mkdir output
mkdir report
```

After a series of commands in the terminal (initializing, configuring, etc.), I have pushed the necessary files into my [repository](https://github.com/epm927/Workout-1.git). 

### Downloading the Data:

Using the `curl -O` command (in the terminal, inside the `data/` folder), I downloaded the data.


\ 

## 2) Create a **`README`** file

I added some text to the README file, outlining the goals of this project as well as dating the project. 

\ 

## 3) Create a Data Dictionary

I used the command `touch nba2018-dictionary.md` (while inside the `data/` directory) to create a blank data dictionary. Using `open`, I opened the file (using LaTeX, by default) and edited the dictionary. To examine the data, I will import it here into `R`, and perform exploration:

```{r}
dat <- read.csv("../data/nba2018.csv")
length(dat[,1])
```

Looks like there are 477 lines of data. After adding this information to the dictionary, I added, committed, and pushed everything onto the repository again.

\ 

Now I switch to an `R` script. See you on the flipside!